Stellar Radiation is Critical for Regulating Star Formation and Driving Outflows in Low Mass Dwarf Galaxies

### 2019-03-15 thesis/chapters/ch_introduction.tex
update to onezone section and Leo P section

%
% 
%
\textbf{Not quite sure if onez models should be its own section:} One of the most common ways to model and understand the stellar populations of galaxies in high-dimensional chemical space is through simplified one-zone (or many-zone) models. This treatment extends back many decades \citep[e.g.][]{Schmidt1963,TalbotArnett1971,Lynden-Bell1975}, where galaxies and their total metal abundance were treated as closed-box systems of gas and stars. Over the interveaning decades these models have improved, including perscriptions for gas outflow (a ``leaky-box") or gas inflow (``accreting box"), or both (a ``bathtub" model, e.g. \cite{FinlatorDave2008,Bouche2010}). The assumptions that go into these models vary dramatically with complexity. Simple models can assume, for example, that long-lived low mass stars do not contribute to metal enrichment while metals from massive stars are instantaneously recylced into the galaxy's gas reservoir (i.e. ignoring stellar lifetimes) or that all metals mix instantly and homogenously. More complex models account for individual stellar lifetimes (to some degree), or may try and account for some aspects of inhomogenous mixing by constructing mulit-zone models that may separate a galaxy in radial bins (with or without mixing between bins), may account for a hot and cold component separately, with mixing between the two, or account for separate ISM and CGM components. The output of these models is typically the mean metal abundance evolution within each of the tracked components. There is generally no self-consistent perscription to account for the scatter in any abundance relationship in these models, which, when needed, is often added ad-hoc in post-processing. Regardless of the assumptions that go into these models, they are powerful tools that can be used to probe the general physical processes that govern galactic chemical evolution, and explore the vast, uncertain parameter space associated with the models that characterize these processes \citep[e.g.][]{Cote2017a}.

@@

\section{Ingredients of Galactic Chemical Evolution}

@@

We dedicate this section to a specific Local Group dwarf galaxy, Leo P, as its observed properties motivate the initial conditions of the simulations developed in this dissertation. This galaxy is remarkable for being one of the lowest mass dwarf galaxies with observed, ongoing star formation. It is located at a distance of $\sim$ XXX (cite), and was first discovered in HI (cite). Since then, multiple studies have examined its star formation history (cite), gas phase metal abundance as traced through emission in an HII region (cite), stellar abundances (cite), and, most recently, resolved spectroscopy of both the gas and stars

%
%
%
%\section{Semi-Analytic Models / One-zone Models}
%
%Definitely have a lot here. Motivate need for better understanding of mixing processes in galaxies. \citep{SchonrichWeinberg2019}
%
%

>>>>

@@

\section{Ingredients of Galactic Chemical Evolution}
\label{intro:sec:ingredients}

@@

We dedicate this section to a specific Local Group dwarf galaxy, Leo P, as its observed properties motivate the initial conditions of the simulations developed in this dissertation. This galaxy is remarkable for being one of the lowest mass dwarf galaxies with observed, ongoing star formation. It is located at a distance of $1.62\pm 0.15$~Mpc \citep{McQuinn2015a}, and was first characterized as part of the ALFALFA survey \citep{Giovanelli2013,Rhode2013,Skillman2013,McQuinn2013,Bernstein-Cooper2014}. These studies have characterized its star formation history as low and roughly continuous over the age of the Universe (value, cite), its stellar mass ($M_{*} = 5.7 \times 10^{5}$~M$_{\odot}$, \cite{McQuinn2013}), its HI and dynamical mass ($M_{\rm HI}(r<r_{\rm HI}) = 9.5\times 10^{5}$~M$_{\odot}$, $M_{\rm dyn}(r<r_{\rm HI}) = 2.6 \times 10^{7}$~M$_{\odot}$, $r_{\rm HI} = $, \cite{Bernstein-Cooper2014}), and gas-phase abundances from an HII region (12 + log(O/H) = 7.17$\pm$0.04 \cite{Skillman2013}). Finally, we chose to focus on a Leo P like galaxy in our simulations in part to allow for comparisons to the work in \cite{McQuinn2015}, which makes an accounting of the total metals (as traced by O) contained in the ISM and stars in Leo P, estimating the fraction that must have been ejected over the lifetime of the galaxy. This is an important measurement as prior estaimtes of this number for low mass dwarf galaxies were restricted to Milky Way dSph's, which are devoid of gas and contaminated by environmental affects. As of this writing, Leo P is one of the only low mass dwarf galaxies with both gas-phase and stellar metal abundance measurements, but there is ongoing work to greatly expand this sample in the near future.

\section{Onezone Models of Galactic Chemical Evolution}
\label{intro:sec:onezone}

This dissertation utilizes high-resolution hydrodynamics simulations that are expensive to run in both walltime ($\sim$ weeks to months) and computational time ($\sim 10^{5-6}$ CPU hours). For this reason,

 One of the most common ways to model and understand the stellar populations of galaxies in high-dimensional chemical space is through simplified one-zone (or many-zone) models. This treatment extends back many decades \citep[e.g.][]{Schmidt1963,TalbotArnett1971,Lynden-Bell1975}, where galaxies and their total metal abundance were treated as closed-box systems of gas and stars. Over the interveaning decades these models have improved, including perscriptions for gas outflow (a ``leaky-box") or gas inflow (``accreting box"), or both (a ``bathtub" model, e.g. \cite{FinlatorDave2008,Bouche2010}). The assumptions that go into these models vary dramatically with complexity. Simple models can assume, for example, that long-lived low mass stars do not contribute to metal enrichment while metals from massive stars are instantaneously recylced into the galaxy's gas reservoir (i.e. ignoring stellar lifetimes) or that all metals mix instantly and homogenously. More complex models account for individual stellar lifetimes (to some degree), or may try and account for some aspects of inhomogenous mixing by constructing mulit-zone models that may separate a galaxy in radial bins (with or without mixing between bins), may account for a hot and cold component separately, with mixing between the two, or account for separate ISM and CGM components. The output of these models is typically the mean metal abundance evolution within each of the tracked components. There is generally no self-consistent perscription to account for the scatter in any abundance relationship in these models, which, when needed, is often added ad-hoc in post-processing. Regardless of the assumptions that go into these models, they are powerful tools that can be used to probe the general physical processes that govern galactic chemical evolution, and explore the vast, uncertain parameter space associated with the models that characterize these processes \citep[e.g.][]{Cote2017a}.

One final motivation of this dissertation is to improve the assumptions within these models by better charaterizing the process of inhomogenous mixing in the ISM, and the coupling of metals to galactic winds and outflows. This is particularly important if the behavior differs between elements from different nucleosynthetic sites. In addition, this work could be used to account for the spread and higher-order statistics of stellar abundance distributions, which would be a powerful tool for better leveraging the wealth of observations in the Milky Way and Local Group. Yet, as discussed, the current state of the art perscriptions for these processes are ad-hoc, lacking physical motivation. Recent analytic work in \cite{KrumholzTing2018} lays down a framework to evaluate the correlation between individual metals in different galactic environments and predicts that the correlations will differ for different nucleosynthetic sites (e.g. AGB winds vs. SNe). We characterize these differences in detail in Chapter~\ref{ch:chapter3}, laying the groundwork for how one could potentially incorporate this insight into improving these models. We build upong this idea more in Chapter~\ref{ch:chapter4}, and plan to explore the connection between our simulations and analytic models in the future. Recent work by \cite{SchonrichWeinberg2019} is a great example of how this could greatly impact current models for galactic chemical evolution. They how adopting a two-phase onezone model (hot and cold ISM), along with parameters for how metals are injected and mix between the phases whose values differ depending on nucleosynthetic origin of the metal, can reconcile long-standing problems in understanding the r-process abundance evolution of stars in the Milky Way. And in fact, they argue that it would be impossible to reproduce observed r-process abundances in the Milky Way (namely [Eu/Fe], [Eu/Si], and [Eu/Mg] as functions of [Fe/H]) with a standard one-phase model.
